# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

The first thing to do is to import the library that we'll need in this analysis.
```{r}
library(ggplot2)
```

We can now start to open the csv file, convert it to a data.frame and preprocess the data in order to have a tidy and useful dataset.
```{r}
#open the dataset
dataset <- read.csv("./activity.csv")
#convertions
dataset$date <- as.Date(dataset$date)
dataset$steps <- as.numeric(as.character(dataset$steps))
dataset$interval <- as.numeric(as.character(dataset$interval))
head(dataset)
#preprocessing
steps_sum <- as.data.frame(tapply(dataset$steps, dataset$date, sum, na.rm=TRUE))
colnames(steps_sum)[1]<- "steps"
``` 
We now have a tidy and simple dataset to analize:
```{r}
str(steps_sum)
head(steps_sum)
```


## What is mean total number of steps taken per day?
We can now create a histogram that plot the steps per day.
```{r}
ggplot(steps_sum, aes(x = as.Date(row.names(steps_sum)), y = steps_sum$steps)) +
  geom_histogram(stat="identity") + labs(title = "Steps Sum", y = "Steps", x = "Days") +
  theme(plot.title = element_text(size = 20, lineheight=.8, face="bold"))
```
We can proceed calculating the mean of the steps:
```{r}
steps_mean <- mean(steps_sum$steps)
steps_mean
```
And the median of the steps:
```{r}
steps_median <- median(steps_sum$steps)
steps_median
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
